---
title: "hw-1"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

1.  (20 points) First, let’s look at the distribution of data (Yi = number correct) we would expect to see if θi = 0.2 across the whole population, i.e. everyone is at chance performance (no ESP). Using the R functions ggplot (with geom point and geom line) and dbinom, plot this distribution.

```{r}
n <- 25  # total number of trials
theta <- 0.2  # probability of success

Yi <- 0:n  # possible outcomes (0 to n)
prob <- dbinom(Yi, size = n, prob = theta)

# create a data frame for plotting
data <- data.frame(Yi = Yi, Probability = prob)

# Plot the distribution using ggplot
ggplot(data, aes(x = Yi, y = Probability)) +
  geom_point(color = "blue", size = 3) +  
  geom_line(color = "blue", size = 1) +   
  labs(title = "Binomial Distribution (n = 10, θ = 0.2)",
       x = "Number Correct (Yi)",
       y = "Probability") +
  theme_minimal()

```

2.  (20 points) Simulate data (using the R function rbinom) from n = 1000 people under the same assumption that θi = 1/5 = 0.2 (i.e. chance performance/no ESP). Make a histogram of this data. It should have the same shape as the binomial probability distribution you plotted in the previous question.

```{r}
# Parameters
n_people <- 1000  # Number of people
n_trials <- 10    # Number of trials for each person
theta <- 0.2      # Probability of success

# Simulate data using rbinom
simulated_data <- rbinom(n = n_people, size = n_trials, prob = theta)

# Create a data frame for plotting
data <- data.frame(Yi = simulated_data)

# Plot the histogram using ggplot
ggplot(data, aes(x = Yi)) +
  geom_histogram(binwidth = 1, color = "black", fill = "blue", alpha = 0.7) +
  labs(title = "Histogram of Simulated Data (n = 1000, θ = 0.2)",
       x = "Number Correct (Yi)",
       y = "Frequency") +
  theme_minimal()

```

3.  (20 points) Repeat question 1, but now assume that our population has psychic powers, so everyone has a 90% chance of guessing the correct card on each trial.

4.  (20 points) Repeat question 2, but continue to assume that everyone has ESP (90% chance of guessing correctly).

5.  Now we will do a third simulation where we assume that 95% of the population has no ESP (i.e. θi = 0.2), but the remaining 5% of people have ESP (with a 90% chance of guessing correctly on each trial).
