---
title: "PSYC 529 R Homework 4"
author: 
- name: "Jacqueline Heitmann"
  affiliation: "25562334"
format: 
  typst: 
    section-numbering: 1.1.a
    keep-typ: true
    include-before-body: 
    - text: |
        #show heading: head => [
          #pagebreak(weak: true)
          #head
          ]
---
```{r}
library(tidyverse)
words <- read.csv("../data/word_memory_data_summarized.csv")
words
```

1. Looking at the data table (you don’t need to import it as an R data frame). Define the following
as variables in R:
(a) (21/2 points) N: the number of relevant data points (trials)
```{r}
memorized_ds <- words %>%
  filter(condition == "Memorized")

N <- sum(memorized_ds$trials)

N
```

(b) (21/2 points) z: the number of “yes” responses on relevant trials
```{r}
z <- sum(memorized_ds$yes_responses)
z
```

(c) (21/2 points) a: the first prior beta hyperparameter
(d) (21/2 points) b: the second prior beta hyperparameter
```{r}
# beta(4,4)
a <- 4
b <- 4
```


2. (10 points) Use these variables to compute the posterior beta distribution hyperparameters
(call them a post and b post).

```{r}
a_post <- a + z # prior + successes
b_post <- b + N - z # prior + failures

a_post
b_post
```
3. We now make a combined plot of the prior and posterior distributions. Hint: Use the file “beta
conjugate prior demo” on Canvas as a template for doing all of this. That will make this part
of the homework much easier and faster

data frame for plotting prior distribution:
```{r}
prior_plot_df = data.frame(theta = seq(from = 0, to = 1, length.out = 200),
                           distribution = "prior") %>%
  mutate(p_theta = dbeta(theta, shape1 = a, shape2 = b))
```

data frame for plotting the posterior distribution:
```{r}
post_plot_df = data.frame(theta = seq(from = 0, to = 1, length.out = 200),
                          distribution = "posterior") %>%
  mutate(p_theta = dbeta(theta, shape1 = a_post, shape2 = b_post))
```

combine prior and posterior plotting data frames:
```{r}
combo_plot_df = bind_rows(prior_plot_df,
                          post_plot_df)
```

combined plot:
```{r}
combo_plot_df %>% ggplot(aes(x = theta, y = p_theta, color = distribution)) +
  geom_line() +
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) # more marks on x axis
```
4. The expected value (mean) and mode (most probable value) are two ways to summarize a
probability distribution. Given the hyperparameters a and b of a beta distribution, you can
compute the distribution’s mode (ω) and expected value (i.e. mean, µ) using formulas given in
the textbook (page 129) and summarized in Cliff’s notes for chapter 6 (these are on Canvas).
Use these formulas and your R variables to compute and display the following:

(a) (10 points) E[θ] (the prior mean of θ)
```{r}
e_prior <- a / a + b
e_prior
```
(b) (10 points) the prior mode of θ.
```{r}
mode_prior <- (a - 1) / (a + b - 2)
mode_prior
```

(c) (10 points) E[θ|N, z] (the posterior mean of θ)

```{r}
e_post <- a_post / (a_post + b_post)
e_post
```

(d) (10 points) the posterior mode of θ
```{r}
mode_post <- (a_post - 1) / (a_post + b_post - 2)
mode_post
```

5. You can use the R function qbeta to find quantiles of a beta distribution. 

For example, let’s suppose we want to find the median (0.5 quantile, i.e. 50th percentile) of a beta(3, 2) distribution.
Then we use the following code: qbeta(p = 0.5, shape1 = 3, shape2 = 2)

This tells use that, assuming a = 3 and b = 2, we have p(θ ≤ 0.6142724) = 0.5. In other words
the median of a beta(3, 2) distribution is about 0.61. 

Use qbeta to compute the following:
(a) (5 points) The 5th percentile (i.e. 0.05 quantile) of the prior distribution
```{r}
qbeta(p = 0.05, a, b)
```
the probability of theta being less than or equal to 0.2253216 is roughly 5% in the prior distribution.

(b) (5 points) The 5th percentile of the posterior distribution
```{r}
qbeta(p = 0.05, a_post, b_post)
```
The probability of theta being less than or equal to 0.78 is roughly 0.05. 

(c) (5 points) The 95th percentile (i.e. 0.95 quantile) of the prior distribution
```{r}
qbeta(p = 0.95, a, b)
```

(d) (5 points) The 95th percentile of the posterior distribution
```{r}
qbeta(p = 0.95, a_post, b_post)
```

Note: We can use these percentiles to construct 90% credible intervals for θ: the probability
that θ is between the 5th and 95th percentiles of the distribution is 90%. For the sake of time
we won’t go into this in detail, but ask me if you have any questions about it (it’s very similar to
how you derive frequentist confidence intervals, except using the prior or posterior distribution
over θ instead of the sampling distribution of some estimator).



