---
title: "Simple Stan Demo (Bernoulli-beta model)"
output: html_notebook
---

```{r}
library(tidyverse) # for general use
library(rstan) # for interacting with Stan
library(bayesplot) # for plotting
```
#### Import data and select my data
```{r}
Zener = read.csv("../data/Zener_data.csv") # import the data
my_Zener = Zener %>% filter(id == "spp36") # select my data
my_Zener %>% head() # show the first few rows
```

#### Count number correct
```{r}
sum(my_Zener$correct)
```

#### Make a list of data (plus beta prior hyperparameters)
```{r}
stan_data_list = list(N = nrow(my_Zener), # number of trials (i.e. rows in the data frame)
                      y = my_Zener$correct) # 1 = correct, 0 = wrong
```

#### Visualize the prior distribution (similar code to last week)
```{r}
data.frame(theta = seq(from = 0, to = 1, length.out = 200),
                           distribution = "prior") %>%
  mutate(p_theta = dbeta(theta, shape1 = 10.6, shape2 = 39.4)) %>%
  ggplot(aes(x = theta, y = p_theta)) + 
  geom_line(color = "blue") +
  xlim(c(0, 1)) +
  theme_classic()
```

#### Fit the model, i.e. do MCMC sampling
```{r}
bern_beta_fit = stan(file = "../Stan files/Bernoulli_beta.stan",
                     model_name = "basic_Zener_card_model",
                     data = stan_data_list,
                     chains = 4,
                     iter = 5000,
                     warmup = 200)
```

#### Trace plot of theta (uses the bayesplot package)
```{r}
mcmc_trace(bern_beta_fit,
           pars = c("theta"))
```

#### Print model summary
```{r}
print(bern_beta_fit, pars = c("theta"))
```

#### Histogram of posterior distribution (uses the bayesplot package)
```{r}
mcmc_hist(bern_beta_fit,
          pars = c("theta")) +
  xlim(c(0, 1)) # make sure that the x-axis shows the full range of possible theta values (0 to 1)
```

#### Plot posterior distribution with credible interval (uses the bayesplot package)
```{r}
mcmc_areas(bern_beta_fit,
           pars = c("theta"),
           prob = 0.95) + # 95% credible interval
  xlim(c(0, 1)) # make sure that the x-axis shows the full range of possible theta values (0 to 1)
```
#### Plot credible interval for theta (uses the bayesplot package)
```{r}
mcmc_intervals(bern_beta_fit, 
               pars = c("theta"),
               prob = 0.5, # inner 50% credible interval
               prob_outer = 0.95) # outer 95% credible interval
```
