---
title: "Homework 5"
author:
- name: "<YOUR NAME>"
  affiliation: "<YOUR UPI>"
format:
  typst:
    section-numbering: 1.1.a
    keep-typ: true
    margin: { x: 0.5in, y: 0.5in }
    include-before-body:
    - text: |
        #show heading: head => [
          #pagebreak(weak: true)
          #head
        ]
---

# Question 1A

```{r q1a}
```

# Question 1B

```{r q1b}
```

# Question 1C

```{r q1c}
```

# Question 1D

```{r q1d}
```

# Question 1E

```{r q1e}
```

# Question 1F

```{r q1f}
```

# Question 1's Stan Code
```{r q1stan, echo=F, results="asis"}
q1.stan <- readLines("question1.stan")
cat("```{.stan filename=\"question1.stan\" code-line-numbers=\"true\"}\n")
cat(paste(q1.stan, collapse = "\n"))
cat("\n```\n")
```

# Question 2A

```{r q2a}
```

# Question 2B

```{r q2b}
```

# Question 2C

```{r q2c}
```

# Question 2D

```{r q2d}
```

# Question 2's Stan Code
```{r q2stan, echo=F, results="asis"}
q2.stan <- readLines("question2.stan")
cat("```{.stan filename=\"question2.stan\" code-line-numbers=\"true\"}\n")
cat(paste(q2.stan, collapse = "\n"))
cat("\n```\n")
```

# Question 3A

```{r q3a}
```

# Question 3B

```{r q3b}
```

# Question 3C

```{r q3c}
```

# Question 3D

```{r q3d}
```

# Question 3E

```{r q3e}
```

# Question 3's Stan Code
```{r q3stan, echo=F, results="asis"}
q3.stan <- readLines("question3.stan")
cat("```{.stan filename=\"question3.stan\" code-line-numbers=\"true\"}\n")
cat(paste(q3.stan, collapse = "\n"))
cat("\n```\n")
```
